\documentclass[12pt]{article}
\input{../preamble.tex}

\begin{document}
\centerline {\textsf{\textbf{\LARGE{Homework 3}}}}
\centerline {Jaden Wang}
\vspace{.15in}

\begin{problem}[6.5]
No. Since $ f: \rr \to \rr, x \mapsto \sqrt{x^2+1}$, notice $ \sqrt{x^2+1} > x, \sqrt{y^2+1} >y$, and $ x-y$ and  $ \sqrt{x^2+1} + \sqrt{y^2+1}$ has the same sign so we can drop the absolute value, then
\begin{align*}
	\frac{(x-y)}{ \sqrt{x^2+1} - \sqrt{y^2+1} } &= \frac{(x-y)(\sqrt{x^2+1}+ \sqrt{y^2+1})}{ x^2+1 - y^2 - 1} \\
	&= \frac{\sqrt{x^2+1} + \sqrt{y^2+1}}{ x+y} \\
	&> 1
\end{align*}
So we establish that $| f(x) - f(y) | < |x-y|$. However, suppose $\sqrt{x^2+1}  = x $ this forces $ 0=1$ a contradiction so  $ f$ has no fix-point. Since  $ \rr$ is a complete metric space, we have a counterexample.
\end{problem}

\begin{problem}[6.7]
	Define $ \gamma: [0,1] \to \rr^{n}, t \mapsto (1-t)p+tq$. Then $ f \circ \gamma: [0,1] \to \rr$. Notice that $ \gamma'(t) = q-p$. Applying the 1D mean value theorem to this function yields a $ t \in (0,1)$ s.t.\ 
	\begin{align*}
		\frac{f \circ \gamma (1) - g \circ \gamma (0) }{ 1-0} &= (f \circ \gamma)'(t)\\
		f(q) - f(p)&= Df(s) \circ \gamma'(t) & s:= (1-t)p+tq \\
		f(q) - f(p) &= Df(s) (q-p).
	\end{align*}
\end{problem}

\begin{problem}[6.10]
\begin{thm}
Let $ f: \rr \to \rr$ be a smooth map. Suppose that $ f'$ is nonzero at some $ p \in \rr$. Then $ f^{-1}$ exists in some open interval around $ p$ and is also smooth. Moreover,
\begin{align*}
	(f^{-1})'(f(p)) = \frac{1}{f'(p)}.
\end{align*}
\end{thm}
\begin{proof}
Since $ f'(p)$ is nonzero, it is either positive or negative. WLOG suppose $ f'(p)>0$, then since $ f'$ is continuous, $ f$ is monotone on  some interval $ (c, d)$ containing $ p$. Hence $ f^{-1}$ exists on this interval. Smoothness of $ f ^{-1}$ on $ (c,d)$ is the result that $ f \circ f ^{-1} = \text{id} $ and $ f$ and  $ \text{id} $ are both smooth so $ f^{-1}$ must be smooth. By chain rule,
\begin{align*}
	(f \circ f^{-1})'(f(p)) &= \text{id}'(f(p))  \\
	f'(p) (f^{-1})'(f(p))&= 1 \\
	(f^{-1})'(f(p)) &= \frac{1}{f'(p)} 
\end{align*}
\end{proof}
\end{problem}

\begin{problem}[6.12]
Suppose the theorem is true when $ M = \rr^{m}$ and $ N = \rr^{n}$. Then for a smooth map of general manifolds (with dim $ m,n$) $ f:M \to N$, take any local charts $ (U,\phi),(V,\psi)$ of $ p$ and  $ f(p)$, we see that $ \psi \circ f \circ \phi ^{-1}: \rr^{m} \to \rr^n$ by assumption yields local charts (linear isomorphisms) $(U', \phi')$ and $ (V', \psi')$ for $ \rr^{m}$ and $ \rr^{n}$ respectively s.t.\ 
\begin{align*}
	\psi' \circ (\psi \circ f \circ \phi^{-1}) \circ \phi' ^{-1}(x_1,\ldots,x_n) &= \Psi \circ f \circ \Phi^{-1} (x_1,\ldots,x_n)\\
	&= (x_1,\ldots,x_k,0\ldots,0) 
\end{align*}
Then $\Phi := \phi' \circ  \phi$ and $\Psi:= \psi' \circ \psi$ are the local charts we seek for $ f$.

Now assume $ f: \rr^{m} \to \rr^{n}$. If $ p \neq 0$, then we can set  $ \wh{ f}(x) = f(x-p) $ which has the same Jacobian as $ f$ so it doesn't change the proof. Similarly, if $ f(p) \neq 0$, we can set  $ \widetilde{ f}(x) = f(x) - f(p)$ and again it doesn't change the Jacobian. Finally, by assumption $\rank Df =k$, so  $ Df(0)$ has  $ k$ linearly-independent columns and $ k$ linearly independent rows. So by a series of permutation matrices we can swap all the linearly-independent columns to the first $ k$-columns, and then swap all the linearly-independent rows to the first $ k$-columns. Then the first $ k \times k$ submatrix has $ k$ pivots so it is nonsingular. These permutations are nonsingular and smooth which still allow us to use the inverse function theorem.
\end{problem}

\begin{problem}[6.13]
Suppose $ f: \rr^2 \to \rr$ is $ C^{1}$ and 1-1. Then $ \rank d_pf \leq \dim \rr = 1$ so it is either 0 or 1. Suppose $ \rank d_pf = 0$ for all $ p \in \rr^2$, then $ f$ is clearly not 1-1  as it is constant. Suppose there exists a $ p \in \rr^2$ s.t.\ $ \rank d_pf = 1$. Then we know that  $ f(p)$ is a regular value by definition, but since  $ f$ is 1-1, the preimage $f^{-1}(f(p)) = \{p\} $ which is a submanifold. But $ p$ is clearly not a manifold, a contradiction. Hence in both cases no such function can exist.
\end{problem}

\begin{problem}[1.3.5]
Recall that $ f:X \to Y$ is a local diffeomorphism if for all $ x \in X$, $ f$  maps a neighborhood of $ x$ diffeomorphically to a neighborhood of  $ f(x)$. 

Since $ f$ is clearly surjective onto its image, and $ f$ is 1-1 by assumption, we have  $ f:X \to f(X)$ is a bijection so $ f^{-1}$ is well-defined as a set map. Denote each neighborhood of $ x$ locally diffeomorphic to open set of $ Y$ as  $ U_x$. Let $ V_y = f(U_x)$ with  $ y = f(x)$. Then each $ V_y$ is open by local diffeomorphism so $ W:=\bigcup_{ x \in X} V_y $ is an open subset of $ Y$. Since $ W \subseteq f(X)$ but $ f(x) \in W \ \forall \ x \in X$, we have $ W = f(X)$. Moreover, $ (f^{-1})|_{V_y} = (f|_{U_x})^{-1}$ so $ f^{-1}$ is locally smooth as well. Smoothness of $ f$ follows from that $ f$ is locally smooth so any transition map restricted to such neighborhood is smooth. These restricted neighborhood for all points in $ X$ still form an atlas of $ X$ so  any transition map from this atlas is smooth. Likewise $ f^{-1}: W \to X$ is smooth. Therefore, $ f$ is a diffeomorphism from  $ X$ to  $ W$.
\end{problem}
\end{document}
