\documentclass[12pt,class=article,crop=false]{standalone} 
\input{../preamble.tex}

\begin{document}
\section{Tangent Spaces}
There is no linear structure on manifolds.

\begin{eg}[tangent space on $ \rr^n$]
$ T_p \rr^{n} = \{(p,q): q \in \rr^{n}\}$. This is a vector space. Note that $ T_p \rr^{n} \cong \rr^{n}, (p,q) \mapsto q-p$. This is canonical. We lose this canonical isomorphism in other manifolds.
\end{eg}

From the submanifold perspective:
\begin{claim}
$ \dim(T_pM) = n $.
\end{claim}
\begin{proof}
\begin{align*}
	\alpha &= \phi ^{-1} \circ  (\phi \circ  \alpha)\\
	\alpha'(0)&= D_0 \phi ^{-1} (\phi \circ \alpha)'(0) 
\end{align*}
\begin{claim}
$ \rank D_0 \phi ^{-1} = n$.
\end{claim}
Since $ \overline{\phi} \circ \phi ^{-1} = \text{id}_{ \phi(U)} $, by chain rule,
\begin{align*}
	D \overline{\phi} \circ D \phi ^{-1} = I
\end{align*}
which yields that $ \rank D \overline{\phi} \circ D \phi ^{-1} =  n$. It follows that $ \rank D \phi_^{-1} \geq n$ which yields equality.
\end{proof}

From the intrinsic perspective, we define $ d f_p: T_pM \to T_{f(p)}N$. Given $  v \in T_pM$, where $ v = \alpha'(0)$ for some curve $ \alpha$ that passes through $ p$. Then define
 \begin{align*}
	d f_p (v) := (f \circ \alpha)'(0)
\end{align*}
Exercise: If $ M = \rr^{n}$ and $ N = \rr^{m}$, then $ d f_p(v) $ reduces to the Jacobian at  $ p$ multiplying by  $ v$ (directional derivative).

Define $ \gamma'(0)$ by operating on functions $ M \to \rr$:
\begin{align*}
	[\gamma'(0)](f) := (f \circ \gamma)'(0)
\end{align*}
So $ \gamma \sim \widetilde{ \gamma}$ iff $ [ \gamma(0)] f = [ \gamma'(0)]f$ for all $ f \in \mathcal{ F}(M)$ space of smooth functions on $ M$.

 \begin{prop}
$ T_pM$ is a vector space with natural operations and has the same dimension as that of the manifold.
\end{prop}
\begin{proof}
\begin{align*}
	[ \alpha'(0)] f &= (f \circ \alpha)'(0) \\
	&= (f \circ x \circ x^{-1} \circ \alpha)'(0) \\
	&= \sum_{ i= 1}^{ n} \frac{ \partial (f \circ x)}{ \partial x_i} ((x^{-1} \circ \alpha)^{i} )'(0)
\end{align*}
Define local coordinates via $ x: \rr^{n} \to M. x_i(t) = x(0,\ldots,0,t,0,\ldots0)$ at $ i$th entry. Then
 \begin{align*}
	 [x_i'(0)] f = (f \circ  x_i)'(0) = \frac{ \partial (f \circ x)}{ \partial x_i} (0)= \sum \lambda_i [x_i'(0)](f).
\end{align*}
So $ \alpha'(0)$ is a linear combination of $ (x_i)'(0)$. It remains to show they are linearly independent. Suppose $ \sum \lamdba_i x_i'(0) =0$.
\end{proof}

\begin{defn}
Let $ f: M \to N$, $ p \in M$, define $ df_p ( \alpha'(0)) = (f \circ \alpha)'(0)$.
\end{defn}
Exercise: this is linear. If $ M = \rr^{n}$, $ N = \rr^{m}$, $ df_p$ is the Jacobian.

\begin{thm}[Chain Rule]
Let $ f: M \to N$ and $ g: N \to L$. Then
\begin{align*}
	d(g \circ f) _p = dg_{f(p)} \circ df_p.
\end{align*}
\end{thm}
\begin{proof}
\begin{align*}
	d(g \circ f) ( \alpha'(0)) &= (g \circ f \circ \alpha)'(0) \\
	&= dg( (f\circ \alpha)'(0)) \\
	&= dg \circ df( \alpha'(0)) 
\end{align*}
\end{proof}

\begin{coro}
If $ M,N$ are diffeomorphic, then  $ \dim M = \dim N$.
\end{coro}
\begin{proof}
Let $ f:M \to N$ be the diffeomorphism, then
\begin{align*}
	f \circ f^{-1} &= \text{id}_M  \\
	df \circ df^{-1} &= I 
\end{align*}
Thus $ df$ is invertible and  $\dim M = \dim T_pM = \dim T_{f(p)}N = \dim N$.
\end{proof}
\end{document}
