\documentclass[12pt,class=article,crop=false]{standalone} 
\input{../preamble.tex}

\begin{document}
\section{Transversality}

Motivation: we want to prove the following theorem but have trouble at tangency.
\begin{thm}[Whitney]
If $ X \subseteq \rr^{n}$ is any closed set, there exists a smooth function $ f: \rr^{n} \to \rr$ with $ X= f^{-1}(0)$. Then the graph $ M = \{(x,f(x)): x \in \rr^{n}\} $ is a submanifold of $ \rr^{n+1}$. 
\end{thm}

\begin{defn}
Let $ X,Y \subseteq Z$ be submanifolds of $ Z$. We say $ X,Y$ are  \allbold{transversal}, denoted $ X \pitchfork Y$ if
 \begin{align*}
	T_p X + T_p Y = T_p Z \ \forall \ p \in X \cap Y.
\end{align*}
\end{defn}

\begin{remark}
If $ X \cap Y = \O$, then $ X,Y$ are trivially transversal.

If $ \dim X + \dim Y < \dim Z$ and $ X \pitchfork Y$, then  $ X \cap Y = \O$.
\end{remark}

The main results are
\begin{enumerate}[label=(\arabic*)]
	\item $ X \pitchfork Y$ and $ X \cap Y = \O$, then $ X \cap Y$ is a manifold. Moreover, $ \dim (X \cap Y) = \dim X + \dim Y - \dim Z$.
	\item Any pair of submanifold becomes transversal after a perturbation (we say they assume general position).
	\item If $ \dim X + \dim Y = \dim Z$, and $ X \pitchfork Y$, then $ X \cap Y$ is a discrete set. If $ X,Y$ are compact, then  $ \#(X \cap Y)$ is finite, called the  \allbold{intersection number}. This number mod 2 is invariant under homotopy. 
\end{enumerate}

Application: general Jordan curve theorem. 

\begin{defn}
Let $ Y \subseteq Z$, $ f: X \to Z$. Then we say f is \allbold{transversal}  to $ Y$,  $ f \pitchfork Y$, if
\begin{align*}
	df_p(T_p X) + T_{f(p)} Y = T_{f(p)} Z \ \forall \ p \in f^{-1}(Y).
\end{align*}
\end{defn}

Note that $ X$ is a submanifold of  $ Y$ if there exists an embedding $ f: X' \to Y$  s.t.\ $ f(X') = X$, this is an immersion and a homemorphism. Rewrite $ X=X'$ and consider $ i: X \to Z$ the inclusion map. The rank theorem says that 

Locally any submanifold is the inverse image of 0 via a submersion. 

To prove (1), since $ X \cap Y = i ^{-1}(Y)$, it is equivalent to prove the following:
\begin{thm}
If $ f: X \to Z$ is transversal to $ Y \subseteq Z$, then  $ f^{-1}(Y)$ is a manifold.
\end{thm}
\begin{proof}
Locally $ Y = g ^{-1}(0)$ where $ g$ is a submersion. Then
\begin{align*}
	f^{-1}(Y) &= f^{-1}(g^{-1}(0)) \\
	&= (g \circ f)^{-1}(0)
\end{align*}
It suffices to show that 0 is a regular value of $ g \circ f$. Notice that $ \ker dg_{f(p)} = T_{f(p)} Y$.
\begin{align*}
	dg _{f(p)} \circ df_{p}(T_p X) + d g_{f(p)} (T_pY)  &= d g_{f(p)} (T_p Z) \\
	dg _{f(p)} \circ df_{p}(T_p X)   &= d g_{f(p)} (T_p Z) 
\end{align*}
Since $ g$ is a submersion, we show that  $ d_{g \circ f}$ has full rank.
\end{proof}

\end{document}
