\documentclass[12pt,class=article,crop=false]{standalone} 
\input{../preamble.tex}

\begin{document}
\section{Characteristic Classes}
There is another way to think of Steifel-Whitney classes:
\begin{thm}
There exists a unique function $ w_i: \text{Vect}(M) \to H^{i}(M;\zz /2 )$, where $ \text{Vect}(M) $ is the set of vector bundles over $ M$, for all  $ M$ and $ i$ satisfying
 \begin{enumerate}[label=(\arabic*)]
	\item $ w_i(f^* E) = f^* (w_i(E)) \ \forall \ f: M \to N$
	\item $ w_0(E) =1, w_1(E) = 0 \ \forall \ i > $ fiber dimension of $ E$.
	\item  $ w(E_1 \oplus E_2) = w(E_1) \smile w(E_2)$ where $ w(E)=1+w_2(E)+w_2(E)+ \ldots$. $ E_1 \oplus E_2$ has fibers the direct sum of fibers of $ E_1$ and $ E_2$. That is, given $(E,M,\rr^{m}),(E,N,\rr^{n})$ we get $ (E_1 \times E_2,M \times N, \rr^{m} \times \rr^{n})$. If $ M=N$, let $ \Delta: M \to M \times M$ be the diagonal map. Define $ E_1 \oplus E_2 = \Delta^* (E_1 \times E_2)$.
	\item $ w_1( \gamma_n) \neq 0$ where $ \gamma_n$ is the universal line bundle over $ \rr P^{n}$.
\end{enumerate}
\end{thm}
For 4 recall
\begin{align*}
	\gamma_n=\{(\ell,v) \in \rr P^{n}+ \rr^{n+1}: v \in\ell\} .
\end{align*}
Exercise: $ \gamma_n$ is a line bundle over $ \rr P^{n}$.

Recall $ H^{i}(\rr P^{n}; \zz /2) \cong \zz /2 \ \forall \ 0 \leq i \leq n$. So 4) implies  $ w_1( \gamma_n)$ generates $ H^{1}(\rr P^{n}l \zz /2)$.

get $ \gamma$ over $ \rr P^{\infty}$ and 4) implies $ w_i( \gamma) \neq 0$.

Exercise:
\begin{enumerate}[label=(\arabic*)]
	\item $ i: \rr P^{n} \to \rr P^{m}$ we get $ i^* ( \gamma_m) = \gamma_n$. If $ w_1( \gamma_1) \neq 0$ then true for $ \gamma_n \ \forall \ n$.
	\item Show $ \rr P^{1} \cong S^{1}$ and $ \gamma_1$ is infinite Mobius band. This is non-orientable, so from above $ w_1( \gamma_1) \neq 0$.
\end{enumerate}
It remains to prove part 3 and uniqueness in the theorem. First we look at some consequences.

Easy consequences:
\begin{enumerate}[label=(\arabic*)]
	\item If $ E_1 \cong E_2$ then $ w_i(E_1) = w_i(E_2) \ \forall \ i$ by 1.
	\item If $ E$ is a trivial bundle, then  $ w_i(E) = 0 \ \forall \ i>0$. This follows from obstruction theory. Let's check using Theorem 5. 
		If $ E \to M$ is trivial, let $ f: M to \{x_0\} $, then $ f^* (\{x_0\} \times \rr^{n}) = E$. So $ w_i(E) = f^* (w_i(\{x_0\} \times \rr^{n} )) = 0$.
	\item If $ E'$ is a trivial bundle and  $ E$ any vector bundle then
		 \begin{align*}
			w_i(E \oplus E') = w_i(E)
		\end{align*}
		from 3.
\end{enumerate}
Recall Whitney showed that any $ n$-manifold embeds in  $ \rr^{2n}$ and immerses in $ \rr^{2n-1}$.
\begin{thm}
If $ \rr P^{2^{r}}$ is immersed in $ \rr^{2^{r}+k}$ then $ k$ must be at least  $ 2^{r}-1$.
\end{thm}
That is, Whitney's Theorem cannot be improved for all manifolds.

Note: if $ f: M^{n} \to \rr^{k}$ is an embedding, then we have the normal bundle of $ f(M^{n})$ :
\begin{align*}
	\nu(M) = \{v \in T_x \rr^{k}: v \perp T_x M \ \forall \ x \in f(M)\} .
\end{align*}
And $ TM \oplus \nu(M) = T\rr^{k}|_M = M \times \rr^{k}$.

Exercise: show $ \nu(M)$ is well-defined if $ f$ is just an immersion and we still have
 \begin{align*}
	TM \oplus \nu(M) = f^* T\rr^{k} = M \times \rr^{k}.
\end{align*}

\begin{thm}
If $ M$ is the boundary of a compact manifold  $ W$, then all Stiefel-Whitney numbers are zero.
\end{thm}

\begin{remark}
The converse is also true by Thom.
\end{remark}

\begin{defn}
Given two unoriented manifolds $ M_1$ and $ M_2$, we say they are \allbold{unoriented cobordant} if there exists a compact manifold $ W$  s.t.\ $ \partial W = M_1 \cup M_2$. 
\end{defn}

\begin{coro}
Two closed, connected manifolds $ M_1$ and $ M_2$ are unoriented cobordant iff they have the same Stiefel-Whitney numbers. 
\end{coro}


\end{document}
